---
title: |
  \large \textbf{Lab Notebook - Week 4} 
subtitle: |
 Agency Profile Briefer \& Policy Lab 2 \vspace{0.2cm}
author: "Jody Park"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
colorlinks: true
urlcolor: blue
linkcolor: blue
fontsize: 11pt
bibliography: "references.bib"
csl: "apanodoi.csl"
link-citations: true
reference-section-title: References
header-includes:
  - \usepackage[capposition=top]{floatrow}
  - \usepackage{placeins}   
  - \usepackage{dcolumn}
  - \usepackage{booktabs}
  - \usepackage{siunitx}
  - \usepackage{amsmath}
  - \usepackage{enumerate}
  - \usepackage[shortlabels]{enumitem}
  - \usepackage[hang,flushmargin]{footmisc}	
  - \usepackage{makecell}
  - \usepackage{url}
  - \usepackage{longtable}
  - \usepackage{graphicx}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{xcolor}
  - \usepackage[most]{tcolorbox}
  - \newtcolorbox{qbox}{enhanced,breakable,boxrule=0.5pt,arc=2mm,
    colback=gray!10,colframe=black!50,left=6pt,right=6pt,top=6pt,bottom=6pt}
  - \usepackage{fontspec}
  - \usepackage[margin=1in]{geometry}
  - \usepackage{titling}
  - \setlength{\droptitle}{-1in}
  - \usepackage{setspace}
  - \setstretch{1.2} 
---

```{r setup, include = FALSE, message=FALSE, results=FALSE, echo = FALSE}
# R libraries
# data manipulation
library(tidyverse)
library(dplyr)
library(tidyr)
library(haven)
library(readr)
library(caret)
library(modelr)
library(broom)
library(ggplot2)
library(plotly)
library(png)
library(openxlsx) 
library(kableExtra)
library(patchwork)
library(texreg)
library(estimatr)
library(MASS) 
library(Matrix)
library(sandwich)   # vcovHC
library(lmtest)     # coeftest
library(boot)       # bootstrap
library(conflicted)
library(car)
library(cowplot)
library(AER)
library(MatchIt)
library(labelled)

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

# For figures
knitr::opts_chunk$set(
                      fig.width = 9, fig.height = 7, fig.retina = 3,
                      out.width = "100%",
                      cache = FALSE,
                      echo = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      hiline = TRUE
                      )
options(tinytex.verbose = TRUE)

knitr::knit_hooks$set(plot = function (x, options) {
  float_correct <- function(f, y, opts)  {
    if (is.null(opts$regfloat) || opts$regfloat==FALSE)
      paste0(f(y, opts), "\n\n\\FloatBarrier\n")
    else
      f(y, opts)
  }
  if (!is.null(options$out.width) || !is.null(options$out.height) ||
      !is.null(options$out.extra) || options$fig.align != "default" ||
      !is.null(options$fig.subcap)) {
    if (is.null(options$fig.scap))
      options$fig.scap = NA
    return(float_correct(knitr:::hook_plot_tex, x, options))
  }
  return(float_correct(knitr:::hook_plot_md_base, x, options))
})

```

# Goals for this week's lab notebook

- Review my SSA agency profile briefer draft 
  - SSA's mission: "We are committed to helping maintain the basic well-being and protection of the people we serve. We pay benefits to about 68 million people including retirees, children, and surviving spouses. From birth, to marriage, and into retirement, we are there to provide support throughout life's journey."

- Review others' draft and provide some questions


\bigskip

# Week 4 Reading Notes \& Questions

## Social Security Administration (SSA)

- Authority / structure: gives more cohesion? or conflicts?

- Long-term/ short-term budgeting: how does SSA balance the two? 
  - Congress / Presidency / Taxes ... how would they affect SSA's budget allocations??


\bigskip


## Equal Employment Opportunity Commission (EEOC)

- 8 statutes - civil rights act / discrimination acts / DEI matters here

- Q1. With the current administration, is the EEOC less active in enforcing anti-discrimination laws and promoting diversity, equity, and inclusion in the workplace?

- Q2. Authority is distributed among the Chair, Commissioners, and General Counsel. If they disagree about an enforcement priority, where is that disagreement resolved? Does this structure protect expertise or complicate accountability?

- independent commission: outside of the president pressure

\bigskip

## Environmental Protection Agency (EPA)

- Q1. The breifer links staffing reductions to possible declines in enforcement capacity. What evidence would distinguish insufficient resources from a deliberate change in agency shirking?? 

- Q2. Stakeholder map: scientists and the public under “keep informed.”
  - Q2a. How does the EPA communicate with these stakeholders? 
  - Q2b. How do these stakeholders influence the EPA’s decision-making? 

- EPA does not have regulatory powers over their works

- Authority: comes from grants - expand research and development, and then they can regulate the research

- One big beautiful bill: large cuts to the EPA, one of the major cuts to the EPA was the reduction in staffing

\bigskip

## Federal Emergency Management Agency (FEMA)

- Q1. The briefer mentions replacing time-consuming damage assessments with faster, data-based payments. Which judgments would shift away from field staff? 
  - Q1a. How would they weigh faster assistance against the risk of missing local needs?

- Q2. FEMA answers to DHS, the White House, and several congressional committees. When their priorities conflict, how does that affect FEMA’s discretion? How could we identify which actor drove a decision?


\bigskip

## Bureau of Alcohol, Tobacco, Firearms and Explosives (ATF)

- Q1. Limited structural independence vs substantial operational discretion. 
  - Q1a. Which decisions most need expert judgment, and what oversight could check abuse without removing that judgment?

- Q2. VanDerStok example concerns applying statutory categories to newer technology. How would more detailed legislation change the tradeoff between constraining ATF and allowing adaptation?

\newpage

# Policy Lab 2 - Due Oct 18

- Intergovernmental relations: how do agencies interact with other agencies, Congress, and the President?
  - Also at state and local levels, and with private actors (e.g., contractors, nonprofits, and advocacy groups).

- Diagram of 16 relationships
  - 3-4k report 
  - Programs
  - Rules
  - Grants
  - Enforcement
  - Litigation
    - state
    - federal
    - (delegation also happens)
  - Instruments
    - gramts
    - delegation (second tier delegation?)
    - coop. agreements
    - preemption (written into statutory languages; federal law preempts state law...puts ceiling; backstop for state law; floors)
    - federal laws work with the government 

- Read statutory authority
  - can anchor to one program
  - can be interpreted broadly or narrowly
  - pick most interesting one to write about with good reasons choosing it

- Data associated with the program
  - e.g., Fedscope = Federal workforce data: compare overall executive enterprises; [Fedscope data](https://www.opm.gov/data/datasets/)
  - [Plumbook](https://www.opm.gov/about-us/open-government/plum-reporting/plum-data/): 1980-ish
  - [USAspending](https://www.usaspending.gov/): my thoughts...
  - partnership with the public organizations 
  - [USAjobs](https://www.usajobs.gov/): how the job ads look like? qualifications, major duties

- Statute
  - look for keyward seraches:
  - keywords that might be relevant to the program/state
  - find intergovernmental relations from the statue; play with detailed prompts to get the intergovernmental relations from the statute
  - statutory authority: keyword searching
    - reading bills and laws 

- Tools
  - Look for those tools what/how are they doing it? (the list under "Diagram of 16 relationships")

\bigskip

## Proquest Regulatory Insight

- ProQuest Regulatory Insight is a database that provides access to regulatory information, including guidance documents, rules, and regulations from various federal agencies.
  - [ProQuest Regulatory Insight](https://regulatoryinsight-proquest-com.ezaccess.libraries.psu.edu/)

- Click database in the library website

- Search SSA, Social Security Administration

- Guidance documents are important when they work with intergovernmental relations

- Content type
  - Regulatory History - Public Law (1,271)
  - Regulatory History - Executive Order (129)
  - tree nodeAgency Overview: Regulations and Guidance (56)
  - Agency (53)
  - Department (3)
  - Topic Page (22)
  - Code of Federal Regulations (43,801)
  - Federal Register (36,724)
  - Notice (20,677)
  - Rule (9,294)
  - Proposed Rule (6,453)
  - Presidential Document:Executive Order (106)
  - Presidential Document:Proclamation (91)
  - Guidance and Policy Document (20,950)
  - Research/Report (7,448)
  - Bulletin (3,565)
  - Guidance Document (1,715)
  - Memorandum (1,448)
  - Manual/Handbook (1,281)
  - Historic Regulation (659)

- Regulatory history
  - what area I want to dig into further??
  - Good way to play around!

- Right top: related database
  - Congressional
  - Legislative insight
  - Supreme court: all litigations; cases

- Pick a particular policy/program to narrow down

- Legislative insight
  - SSA or Social Security Administration: legislative history
  - Find an interesting act and start from there

- Find a rule applies to all states
  - Did it interact differently with different states? How did the states respond to the rule? Did they comply or resist?

- Advanced search - Legislative insight 
  - SSA *OR* Social Security Administration
  - *AND* SSI *OR* SSDI
  - Interesting findings:
    - PL103-296 Social Security Independence and Program Improvements Act of 1994
    - PL98-460 Social Security Disability Benefits Reform Act of 1984
    - PL115-243 Tribal Social Security Fairness Act of 2018
    - PL115-8 Congressional Disapproval of SSA Rule
    - PL111-142 Social Security Disability Applicants' Access to Professional Representation Act of 2010
    - PL101-508 Omnibus Budget Reconciliation Act of 1990

- Statutory authority: look at legislative insight
  - Pick an act and search it
  - Click on the act and download PDF (legislative history of...)
  - save: Statue at Large
  - save: USCS: US code
    - and pull out those USCS
  
- And go to congressional
  - search the act and find background reports, news.. etc.
  - Adminsitering agency

- Statutory $\rightarrow$ Rules $\rightarrow$ Cases $\rightarrow$ Background documents (congressional hearings?) 
  - get an idea of what the act is about and how it was implemented
  - where is fraction/conflicts happending in that intergovernmental relations?

- Prompt: explain that this is a three columns document

- First look / potential choices 
  - I put my detailed research interests/questions/design to my LLM models
  - And this is what I've got:

```{r ssa-policy-lab2-options, results='asis', echo=FALSE}
# compare potential statutes and programs for the intergovernmental relations lab
ssa_lab2_options <- tibble::tribble(
  ~"Rank / act or program", ~"Intergovernmental relationship", ~"Potential research question", ~"Assessment",
  "1. Social Security Disability Benefits Reform Act of 1984 (P.L. 98-460)", "SSA, state DDS agencies, Congress, and courts; federal standards and state compliance.", "How do conflicts over disability administration bring state and federal administrative expertise into congressional deliberation?", "Best overall fit: continues the agency briefer; substantial legislative and regulatory record.",
  "2. Social Security Disability Amendments of 1980 (P.L. 96-265)", "Federal supervision of state disability determinations, reviews, and corrective intervention.", "How does Congress structure federal control while relying on state agencies to implement disability policy?", "Best companion to the 1984 act: explains the administrative arrangements and review controversy preceding reform.",
  "3. Social Security Amendments of 1972 (P.L. 92-603): SSI state supplementation and administration", "Federal SSI benefits, state supplemental payments, and federal-state administrative agreements.", "How do different administrative arrangements affect coordination, accountability, and information needs across states?", "Best alternative: suitable for a contemporary three-state comparison; narrow the broad act to SSI administration.",
  "4. Ticket to Work and Work Incentives Improvement Act of 1999 (P.L. 106-170)", "SSA, state vocational rehabilitation agencies, and public/private employment networks.", "How do payment arrangements and divided responsibilities shape cooperation and performance reporting?", "Strong fit for delegation and implementation; requires clear boundaries because several agencies and programs are involved.",
  "5. Tribal Social Security Fairness Act of 2018 (P.L. 115-243)", "Direct federal-tribal agreements covering tribal council members.", "How does voluntary coverage balance tribal self-government with federal administrative requirements?", "Distinctive, narrow topic; hearing evidence and data for three tribal governments need further checking."
)
knitr::kable(ssa_lab2_options, format = "latex", booktabs = TRUE,
             longtable = TRUE, escape = TRUE,
             caption = "Potential statutes and programs for Policy Lab 2: ranked research options") %>%
  kable_styling(latex_options = "repeat_header", font_size = 9) %>%
  column_spec(1, width = "1.45in") %>%
  column_spec(2, width = "1.35in") %>%
  column_spec(3, width = "1.6in") %>%
  column_spec(4, width = "1.55in")
```

\newpage

## USAspending

- Loans, grants, contracts 

- Explore data
  - choose the agency
 
## LexusNexus / LexusUnique
- LexusUnique: a database that provides access to legal information, including news articles, court cases, and other legal resources.
  - [LexusUnique](https://www.lexisnexis.com/en-us/products/lexis-advance.page)
  - news articles, courtcases
  - LexusNexus
